\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-09-06 paper.qmd
Added intro section for MNIST data

The MNIST dataset consists of $70000$ grayscale images of handwritten digits [@lecun2010]. @yingfan2021 used this dataset to evaluate how PaCMAP preserves local structures in \gD{}. Because of the large number of images, the focus was specifically on the handwritten digit 1, which exhibits a non-linear structure in \gD{} (@fig-pacmap-author). There are $7877$ images of the digit 1 in the dataset. Additionally, PCA was applied as a preprocessing step and the first $10$ principal components were selected. The objective is to assess whether PaCMAP effectively preserves the non-linear structure in \gD{}.

@@

As a result of visualizing the *model-in-the-data-space*, PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data, as it preserves the non-linear structure present in the $10\text{-}D$ data (@fig-mnist-tri-proj). It also shows some quirks like twisted patterns and long edges (@fig-mnist-tri-proj). Furthermore, the results help to identify anomalies (with large model errors) within and the outside the non-linear structure, displaying different patterns of the digit 1 (@fig-model-error-mnist).

>>>>

The digit 1 of MNIST dataset consists of $7877$ grayscale images of handwritten digits [@lecun2010]. Before further analysis, PCA was used to preprocess the data, where the first $10$ principal components, explaining 83% of the total variation, were selected. The objective is to select the best NLDR layout, representing the structure of the digit 1 data in $10\text{-}D$ with default (hyper-)parameters.

<!-- @yingfan2021 used this dataset to evaluate how PaCMAP preserves local structures in \gD{}. Because of the large number of images, the focus was specifically on the handwritten digit 1, which exhibits a non-linear structure in \gD{} (@fig-pacmap-author). There are $7877$ images of the digit 1 in the dataset. Additionally, PCA was applied as a preprocessing step and the first $10$ principal components were selected. The objective is to assess whether PaCMAP effectively preserves the non-linear structure in \gD{}.  -->

@@

<!-- As a result of visualizing the *model-in-the-data-space*, PaCMAP provides a reasonable \gD{} representation of MNIST digit 1 data, as it preserves the non-linear structure present in the $10\text{-}D$ data (@fig-mnist-tri-proj). It also shows some quirks like twisted patterns and long edges (@fig-mnist-tri-proj). Furthermore, the results help to identify anomalies (with large model errors) within and the outside the non-linear structure, displaying different patterns of the digit 1 (@fig-model-error-mnist).  -->

### 2024-09-06 paper.tex
Added intro section for MNIST data

The MNIST dataset consists of \(70000\) grayscale images of handwritten
digits \citep{lecun2010}. \citet{yingfan2021} used this dataset to
evaluate how PaCMAP preserves local structures in \gD{}. Because of the
large number of images, the focus was specifically on the handwritten
digit 1, which exhibits a non-linear structure in \gD{}
(Figure~\ref{fig-pacmap-author}). There are \(7877\) images of the digit
1 in the dataset. Additionally, PCA was applied as a preprocessing step
and the first \(10\) principal components were selected. The objective
is to assess whether PaCMAP effectively preserves the non-linear
structure in \gD{}.

@@

As a result of visualizing the \emph{model-in-the-data-space}, PaCMAP
provides a reasonable \gD{} representation of MNIST digit 1 data, as it
preserves the non-linear structure present in the \(10\text{-}D\) data
(Figure~\ref{fig-mnist-tri-proj}). It also shows some quirks like
twisted patterns and long edges (Figure~\ref{fig-mnist-tri-proj}).
Furthermore, the results help to identify anomalies (with large model
errors) within and the outside the non-linear structure, displaying
different patterns of the digit 1 (Figure~\ref{fig-model-error-mnist}).

>>>>

The digit 1 of MNIST dataset consists of \(7877\) grayscale images of
handwritten digits \citep{lecun2010}. Before further analysis, PCA was
used to preprocess the data, where the first \(10\) principal
components, explaining 83\% of the total variation, were selected. The
objective is to select the best NLDR layout, representing the structure
of the digit 1 data in \(10\text{-}D\) with default (hyper-)parameters.

@@
